---
title: "Forecasting issues"
author: "Forecast Padawan 2"
date: "November 17, 2016"
output: pdf_document
---

The goal of this experiment is to design the best model to forcaste the number of issue in the per day in the comming two weeks. We think that this could help Open Source organisation to manage there human ressources.

# Load the data

```{r results='hide', message=FALSE, warning=FALSE}
#install.packages('forecast')

library('forecast')
library(knitr)
#load the data frame
repository.csv <- read.csv("time_series/tensorflow_tensorflow_daily.csv")

repository.csv$date = as.POSIXlt(as.Date(repository.csv$date,format='%Y-%m-%d'))
```

# keep the last 12 months

```{r}
to_date <- repository.csv$date[length(repository.csv$date)]
from_date <- to_date
from_date$year <- from_date$year - 1

repository.csv <- subset(repository.csv, date <= to_date & date >= from_date)
```

```{r}
#loading issues and commits into a ts object
issues.ts <- ts(repository.csv$number_of_issues, frequency = 7) 


Acf(issues.ts, lag.max = 14, main = "")

commits.ts <- ts(repository.csv$number_of_commits, frequency = 7) 
pull_requests.ts <- ts(repository.csv$number_of_pull_requests, frequency = 7) 

plot(issues.ts, main = 'TensorFlow Issues', bty = 'l', ylab = 'Number of Issues')
plot(commits.ts, main = 'Commits', bty = 'l', ylab = 'Number of Commits')
```

```{r}
time <- time(issues.ts)

n.sample <- 28
n.valid <- 21


separate.train.test <- function(timeserie, n.valid) {
  time <- time(timeserie)
  n.train <- length(timeserie) - n.valid
  results <- list()
  results$train.ts <- window(timeserie, start=time[1], end=time[n.train])
  results$valid.ts <- window(timeserie, start=time[n.train+1], end=time[n.train+n.valid])
  return(results)
}
# create a matrix of 14 column, each column is a time series create by rolling forward
all.issues <- sapply(0:(n.sample - 1), function(i) return(separate.train.test(window(issues.ts,start=time[1],end=time[length(issues.ts)-i]), n.valid)))
all.commits <- sapply(0:(n.sample - 1), function(i) return(separate.train.test(window(commits.ts,start=time[1],end=time[length(commits.ts)-i]), n.valid)))

issues  <- separate.train.test(issues.ts, n.valid)
commits  <- separate.train.test(commits.ts, n.valid)
```

```{r}
# utility functions
# all.forecast is a matirx of 21(length of validation period) * 14(14 rolling forward)
mean.all.accuracy <- function(all.forecast) {
  Reduce("+", all.forecast['summary',])/length(all.forecast['summary',])
}

plot.all.residuals <- function(all.forecast) {
  plot(1, type="l", main="Residuals", xlim=c(35, 53.3), ylim=c(-40, 40), xlab = 'Week', ylab = 'Errors')
  sapply(1:n.sample, function(i) lines(all.forecast['train', i]$train - all.forecast['fitted', i]$fitted))
  sapply(1:n.sample, function(i) lines(all.forecast['residual',i]$residual, col = 'blue'))
  return(NULL)
}

plot.all.pred <- function(all.forecast) {
  plot(issues.ts, main="Prediction", xlim=c(35, 53.3), xlab = 'Week', ylab = 'Number of Issues')
  if (class(all.forecast['pred',1]$pred) == "forecast") {
    sapply(1:n.sample, function(i) lines(all.forecast['pred',i]$pred$mean, col=rgb(0, 0, 1, 0.5)))
  } else {
    sapply(1:n.sample, function(i) lines(all.forecast['pred',i]$pred, col=rgb(0, 0, 1, 0.5)))
  }
  return(NULL)
}

plot.pred <- function(forecast.with.interval.ts) {
  plot(issues.ts, main="Prediction Interval", xlim=c(35, 53.3), xlab = 'Week', ylab = 'Number of Issues')
  # how to plot shade, why is it not working here...~''
  apply(forecast.with.interval.ts, 2, function(x) lines(x))
  return(NULL)
}

hist.all.residuals <- function(all.forecast) {
  residuals <- sapply(1:n.sample, function(i) as.numeric(all.forecast['residual',i]$residual))
  hist(residuals)
  quantile(residuals,c(0.975,0.90,0.10,0.025))
}

# plot the boxplot of 21 validation period prediction residuals
boxplot.all.residuals <- function(all.forecast) {
  residuals <- sapply(1:n.sample, function(i) as.numeric(all.forecast['residual',i]$residual))
  boxplot(apply(residuals, 1, quantile.helper))
  return (quantile(residuals, c(0.975,0.90,0.10,0.025)))
}

# retrun the vector of qunatile of 0.975, 0.90, 0.10, 0.025
quantile.helper <- function(matrix) {
  return (quantile(matrix, c(0.975, 0.90, 0.10, 0.025)))
}

# get the quantile of each point prediction
get.quantile.of.residuals <- function(all.forecast) {
  residuals <- sapply(1:n.sample, function(i) as.numeric(all.forecast['residual',i]$residual))
  return (apply(residuals, 1, quantile.helper))
}

forecast.confidence <- function(ets.test.model.pred, quantile.of.residuals) {
  forecast.confidence.interval <- apply(quantile.of.residuals, 1, function(a.quantile) return(a.quantile+ets.test.model.pred))
  return(forecast.confidence.interval)
}

forecast.manual.interval <- function(x.train, f.train, f.pred, f.lower, f.upper) {
  mean <- f.pred
  x <- x.train
  residuals <- x.train - f.train
  fitted <- f.train
  level <- c(80, 95)
  lower <- f.lower
  upper <- f.upper
  
  # Construct output list
  output <- list(mean=mean, x=x, residuals=residuals, fitted=fitted, level=level, lower=lower, upper=upper)
  # Return with forecasting class
  return(structure(output, class='forecast'))
}

# to build custom forecast object
forecast.manual <- function(x.train, f.train, f.pred) {
  mean <- f.pred
  x <- x.train
  residuals <- x.train - f.train
  fitted <- f.train
  
  # Construct output list
  output <- list(mean=mean, x=x, residuals=residuals, fitted=fitted)
  # Return with forecasting class
  return(structure(output, class='forecast'))
}
```

# Naive Forecast

## Naive

```{r kable}

naive.forecast <- function(sample) {
  results <- list()
  results$train <- sample$train.ts
  results$valid <- sample$valid.ts
  results$pred <- naive(sample$train.ts, h=n.valid)
  results$fitted <- results$pred$fitted
  results$residual <- sample$valid.ts - results$pred$mean
  results$summary <- accuracy(results$pred, sample$valid.ts)
  return(results)
}

all.naive.forecast <- sapply(1:n.sample, function(i) return(naive.forecast(all.issues[,i])))

kable(mean.all.accuracy(all.naive.forecast))
```
```{r}
plot.all.pred(all.naive.forecast)
```
```{r}
plot.all.residuals(all.naive.forecast)
```
```{r}
hist.all.residuals(all.naive.forecast)
```
```{r}
boxplot.all.residuals(all.naive.forecast)
```


## Seasonal Naive

```{r} 
snaive.forecast <- function(sample) {
  results <- list()
  results$train <- sample$train.ts
  results$valid <- sample$valid.ts
  results$pred <- snaive(sample$train.ts, h=n.valid)
  results$fitted <- results$pred$fitted
  results$residual <- sample$valid.ts - results$pred$mean
  results$summary <- accuracy(results$pred, sample$valid.ts)
  
  return(results)
}

all.snaive.forecast <- sapply(1:n.sample, function(i) return(snaive.forecast(all.issues[,i])))

kable(mean.all.accuracy(all.snaive.forecast))
```
```{r}
plot.all.pred(all.snaive.forecast)
```
```{r}
plot.all.residuals(all.snaive.forecast)
```
```{r}
hist.all.residuals(all.snaive.forecast)
```
```{r}
boxplot.all.residuals(all.snaive.forecast)
```

# Smoothing

## Exponential smoothing ZMM

```{r}
hw.forecast <- function(sample) {
  results <- list()
  results$train <- sample$train.ts
  results$valid <- sample$valid.ts
  results$model <- ets(sample$train.ts, model = "ZMM")
  results$pred <- forecast(results$model, h=n.valid)
  results$fitted <- results$pred$fitted
  results$residual <- sample$valid.ts - results$pred$mean
  results$summary <- accuracy(results$pred, sample$valid.ts)
  return(results)
}


all.hw.forecast <- sapply(1:n.sample, function(i) return(hw.forecast(all.issues[,i])))

kable(mean.all.accuracy(all.hw.forecast))
```
```{r}
plot.all.pred(all.hw.forecast)
```
```{r}
plot.all.residuals(all.hw.forecast)
```
```{r}
hist.all.residuals(all.hw.forecast)
```
```{r}
boxplot.all.residuals(all.hw.forecast)
```

## Double differencing

```{r}
ma.dd.forecast <- function(sample) {
  train.issues.d1 <- diff(sample$train.ts, lag = 1)
  train.issues.d1.d7 <- diff(train.issues.d1, lag = 7)
  
  ma.trailing <- rollmean(train.issues.d1.d7, k = 7, align = "right")
  last.ma <- tail(ma.trailing, 1)
  ma.trailing.pred <- ts(c(ma.trailing, rep(last.ma, n.valid)), start=c(3, 1), frequency = 7)
  
  ma.dd.pred.d1 <- train.issues.d1
  ma.dd.pred <- sample$train.ts
  
  for(i in 1:(n.valid/7)) {
    ma.dd.pred.d1 <- ma.trailing.pred + lag(ma.dd.pred.d1,k = -7)
    ma.dd.pred <- ma.dd.pred.d1 + lag(ma.dd.pred,k = -8)
  }
  
  results <- list()
  results$train <- sample$train.ts
  results$valid <- sample$valid.ts
  
  valid.time <- time(results$valid)
  train.time <- time(results$train)
  
  dd.fitted <- window(ma.dd.pred, start=c(5,3), end=end(train.time), frequency=frequency(train.time))
  dd.pred <- window(ma.dd.pred, start=start(valid.time), end=end(valid.time), frequency=frequency(valid.time))
  
  results$pred <- forecast.manual(window(results$train, start=c(5,3)), dd.fitted, dd.pred)
  results$fitted <- results$pred$fitted

  results$residual <- sample$valid.ts - results$pred$mean
  results$summary <- accuracy(results$pred, sample$valid.ts)
  
  return(results)
}

all.ma.dd.forecast <- sapply(1:n.sample, function(i) return(ma.dd.forecast(all.issues[,i])))

kable(mean.all.accuracy(all.ma.dd.forecast))
plot.all.pred(all.ma.dd.forecast)
plot.all.residuals(all.ma.dd.forecast)
hist.all.residuals(all.ma.dd.forecast)
boxplot.all.residuals(all.ma.dd.forecast)
```

# Regression

## Linear additive regression season

```{r}
regr.add.forecast <- function(sample) {
  results <- list()
  results$train <- sample$train.ts
  results$valid <- sample$valid.ts
  results$model <- tslm(sample$train.ts ~ season + trend)
  results$pred <- forecast(results$model, h=n.valid)
  results$fitted <- results$pred$fitted
  results$residual <- sample$valid.ts - results$pred$mean
  results$summary <- accuracy(results$pred, sample$valid.ts)
  
  return(results)
}

all.regr.add.forecast <- sapply(1:n.sample, function(i) return(regr.add.forecast(all.issues[,i])))

kable(mean.all.accuracy(all.regr.add.forecast))
```
```{r}
plot.all.pred(all.regr.add.forecast)
```
```{r}
plot.all.residuals(all.regr.add.forecast)
```
```{r}
hist.all.residuals(all.regr.add.forecast)
```
```{r}
boxplot.all.residuals(all.regr.add.forecast)
```

## linear multiplicative regression

```{r}
regr.mult.forecast <- function(sample.issues) {
  train.ts <- sample.issues$train.ts
  valid.ts <- sample.issues$valid.ts
  train.lm <- tslm(train.ts ~ season + trend, lambda = 0)
  train.lm.pred <- forecast(train.lm, h=n.valid)
  lm.summary <- accuracy(train.lm.pred, valid.ts)
  
  results <- list()
  results$train <- train.ts
  results$valid <- valid.ts
  results$model <- train.lm
  results$pred <- train.lm.pred
  results$fitted <- train.lm.pred$fitted
  results$residual <- valid.ts - train.lm.pred$mean
  results$summary <- lm.summary
  
  return(results)
}

all.regr.mult.forecast <- sapply(1:n.sample, function(i) return(regr.mult.forecast(all.issues[,i])))

kable(mean.all.accuracy(all.regr.mult.forecast))
```
```{r}
plot.all.pred(all.regr.mult.forecast)
```
```{r}
plot.all.residuals(all.regr.mult.forecast)
```
```{r}
hist.all.residuals(all.regr.mult.forecast)
```
```{r}
boxplot.all.residuals(all.regr.mult.forecast)
```

## Neural Network (repeats = 20, p=1, P=1, size=7)


```{r}
nnetar.forecast <- function(sample) {
  results <- list()
  results$train <- sample$train.ts
  results$valid <- sample$valid.ts
  results$model <- nnetar(sample$train.ts, repeats = 20, p=1, P=1, size=7 )
  results$pred <- forecast(results$model, h=n.valid)
  results$fitted <- results$pred$fitted
  results$residual <- sample$valid.ts - results$pred$mean
  results$summary <- accuracy(results$pred, sample$valid.ts)
  
  return(results)
}

all.nnetar.forecast <- sapply(1:n.sample, function(i) return(nnetar.forecast(all.issues[,i])))

kable(mean.all.accuracy(all.nnetar.forecast))
plot.all.pred(all.nnetar.forecast)
plot.all.residuals(all.nnetar.forecast)
hist.all.residuals(all.nnetar.forecast)
boxplot.all.residuals(all.nnetar.forecast)
```

# External info Numerical using regression model


```{r}
regr.ext.forecast <- function(issues, commits.sample) {
  commits_x <- ts(c(commits.sample$train.ts[1:(length(commits.sample$train.ts) - 1)]), frequency = 7, start = c(1,2))
  issues$train.ts <- window(issues$train.ts, start=c(1,2))
  
  newdata <- data.frame(as.numeric(snaive(commits_x, h=n.valid)$mean))
  colnames(newdata) <- c('commits_x')
  
  results <- list()
  results$train <- issues$train.ts
  results$valid <- issues$valid.ts
  results$model <- tslm(issues$train.ts ~ season + trend + commits_x)
  results$pred <- forecast(results$model, h=n.valid, newdata=newdata)
  results$fitted <- results$pred$fitted
  results$residual <- issues$valid.ts - results$pred$mean
  results$summary <- accuracy(results$pred, issues$valid.ts)
  
  return(results)
}

all.regr.ext.forecast <- sapply(1:n.sample, function(i) return(regr.ext.forecast(all.issues[,i], all.commits[,i])))

kable(mean.all.accuracy(all.regr.ext.forecast))

plot.all.pred(all.regr.ext.forecast)
plot.all.residuals(all.regr.ext.forecast)
hist.all.residuals(all.regr.ext.forecast)
boxplot.all.residuals(all.regr.ext.forecast)
```

# Ensemble (all.regr.mult.forecast[,i], all.hw.forecast[,i])


```{r}
ensemble.forecast <- function(list.of.forecast) {
  results <- list()
  results$train <- list.of.forecast[[1]]$train
  results$valid <- list.of.forecast[[1]]$valid
  
  valid.time <- time(results$valid)
  train.time <- time(results$train)
  
  mean.pred <- ts(
    rowMeans(sapply(list.of.forecast, function(forecast) forecast$pred$mean)), 
    start=start(valid.time), 
    end=end(valid.time), 
    frequency=frequency(valid.time))
  
  mean.fitted <- ts(
    rowMeans(sapply(list.of.forecast, function(forecast) window(forecast$fitted, start=c(5,3)))), 
    start=start(train.time), 
    end=end(train.time), 
    frequency=frequency(train.time))
  
  results$pred <- forecast.manual(window(results$train, start=c(5,3)), mean.fitted, mean.pred)
  
  results$fitted <- results$pred$fitted
  
  results$residual <- results$valid - results$pred$mean
  results$summary <- accuracy(results$pred, results$valid)
  
  return(results)
}

all.ensemble.forecast <- sapply(
  1:n.sample, 
  function(i) return(ensemble.forecast(list(all.regr.mult.forecast[,i], all.hw.forecast[,i])))
)

kable(mean.all.accuracy(all.ensemble.forecast))

Acf(all.ensemble.forecast[,1]$residual, lag.max = 14, main = "")

test <- list()
test$train.ts <- issues.ts
test$valid.ts <- naive(issues.ts, h=n.valid)$mean

test.forecast <- ensemble.forecast(list(regr.mult.forecast(test), hw.forecast(test)))
quantile.of.residuals <- get.quantile.of.residuals(all.ensemble.forecast)
# the prediction intrerval of each time stamp
test.forecast.confidence.interval <- forecast.confidence(test.forecast$pred$mean, quantile.of.residuals)
# convert the prediction interval into time series object
test.forecast.confidence.interval.ts <- ts(test.forecast.confidence.interval, start = c(53, 4), end = c(56, 3), frequency = 7)

forecast.object <- forecast.manual.interval(
  x.train=issues.ts, 
  f.train=test.forecast$fitted, 
  f.pred=test.forecast$pred$mean, 
  f.lower=test.forecast.confidence.interval.ts[,1:2], 
  f.upper= test.forecast.confidence.interval.ts[,3:4])
```
```{r}
plot.all.pred(all.ensemble.forecast)
```
```{r}
plot.all.residuals(all.ensemble.forecast)
```
```{r}
hist.all.residuals(all.ensemble.forecast)
```
```{r}
boxplot.all.residuals(all.ensemble.forecast)
```
```{r}
# plot the prediction on test period with the prediction interval
plot(forecast.object, xlim=c(35, 56), main = 'TensorFlow Forecasted No. of issues')
```
